---
title: "PCA"
subtitle: |
  | Project 6
#author : "LAST NAME First name"
#date: "March 19, 2023"
output: 
  html_document:
    theme: cosmo
    highlight: zenburn
    toc: true
    number_sections: true
    toc_float:
      collapsed: false
header-includes:
  - \usepackage{bm}
  - \newcommand{\E}{\mathbb{E}}
  - \newcommand{\R}{\mathbb{R}}
#bibliography: biblio.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
```

COVID-19 has undoubtedly been the most impactful event of the early 2000s. It has impacted every country and the daily lives of individuals worldwide. Measures such as lock downs and isolation were implemented, and rightfully so, as COVID-19 was recognized as a major killer, claiming over 6.7 million lives globally. This unprecedented situation has prompted statisticians and analysts to focus on the available data.

The pandemic data has garnered significant attention from experts in statistics and analysis. These professionals have delved into the data to understand the magnitude of the situation, assess the effectiveness of implemented measures, and attempt to forecast future trends.

In this report, our analysis focuses on the data provided by the official World Health Organization website for countries in Europe. Our objective is to apply the PCA method, a widely recognized technique for dimensionality reduction. 

# Data Set 

The dataset covers the period from January 3, 2020, to May 24, 2023. It is important to note that our analysis focuses on a specific subset of European countries rather than including all countries in the region. The selected countries are shown in the following list : "Austria", "Belgium" ,      "Bulgaria" ,     "Croatia"    ,   "Cyprus"      ,  "Czechia",       "Denmark"   ,   
"Estonia" ,      "Finland",       "France"   ,     "Germany" ,      "Greece"  ,      "Hungary"  ,     "Iceland"  ,    "Ireland"   ,    "Italy"     ,    "Latvia" ,       "Liechtenstein", "Lithuania",     "Luxembourg" ,   "Malta",        
,"Netherlands" ,  "Norway"   ,     "Poland" ,       "Portugal"  ,    "Romania" ,      "Slovakia"    ,   "Slovenia"     
, "Spain"  , and        "Sweden"  . The data we are examining specifically pertains to the cumulative count of deaths attributed to COVID-19, allowing us to gain insights into the impact of the pandemic on the selected European countries over the specified timeframe.

---

However, it's easy to see that country-by-country data are difficult to compare. Intuitively, the more inhabitants a country has, the greater the number of deaths. To get around this problem, a common solution is to transform the data into a cumulative rate per million inhabitants. This is what we obtain in the following figure: 

```{r}
data <- read.csv("https://covid19.who.int/WHO-COVID-19-global-data.csv")
liste_pays <- c("Austria", "Belgium", "Bulgaria", "Croatia", "Cyprus", "Czechia", "Denmark", "Estonia",       "Finland","France","Germany","Greece","Hungary","Iceland","Ireland"  ,     "Italy" ,        "Latvia" ,       "Liechtenstein", "Lithuania" ,    "Luxembourg"  ,  "Malta", "Netherlands" ,  "Norway" ,       "Poland" ,       "Portugal" ,     "Romania",       "Slovakia"  ,    "Slovenia" , "Spain"     ,    "Sweden"        )

pop <-  c(8978929,  11617623,   6838937 ,  3862305 ,   904705  ,10516707,   5873420 ,  1331796,   5548241,  67871925, 83237124,  10459782 ,  9689010 ,   376248,   5060004 , 59030133 ,  1875757 ,    39308 ,  2805998  ,  645397 ,   520971, 17590672,   5425270,  37654247 , 10352042 , 19042455 ,  5434712  , 2107180,  47432893,  10452326)
data = data[data$Country %in% liste_pays,]
pop_list <- as.vector(matrix(unlist(replicate(1238,pop, simplify = FALSE), recursive = FALSE), byrow=TRUE, ncol=30))

data$Cumulative_deaths <- data$Cumulative_deaths/pop_list*1000000

data <- data[,c(1, 3, 8)]
matrice_data <- matrix(data$Cumulative_deaths, byrow=TRUE, ncol=1238)
```

```{r}
max <- max(data$Cumulative_deaths)
min <- min(data$Cumulative_deaths)
plot(1:1238, colMeans(matrice_data),type="l", ylim=c(min, max), col="red")
for (country in 1:30){
  points(1:1238, matrice_data[country,],type="l")
}
points(1:1238, data[data$Country=="Bulgaria",]$Cumulative_deaths, type="l", col=3)
points(1:1238, data[data$Country=="Iceland",]$Cumulative_deaths, type="l", col=2)
```
This representation of the data set allows us to make several observations. 
The first is that as time goes by, the variance of the data increases, and each country deviates more and more from the mean. 
We can also observe a succession of "small waves" which probably show the various waves of the covid before and after confinement. 

It is also worth noting that Bulgaria had the highest number of deaths per capita at the end of the period, while Iceland was the least affected country in Europe. This may make sense, given that Iceland is an island and therefore better protected from the epidemic. What's more, because the island is so vast for so few inhabitants, the population is less concentrated than in some countries, which is favorable in the event of an epidemic such as COVID. 

---

For a better distribution of data, we work with the data log. As shown in the figure below. 

```{r}


data$Cumulative_deaths = log(data$Cumulative_deaths +0.001)

#min(data$cumulative_count[!is.na(data$cumulative_count) & !data$cumulative_count==0])
#data[data$Cumulative_deaths==0 & !is.na(data$Cumulative_deaths),]$Cumulative_deaths = rep(-100, length(data[data$Cumulative_deaths==0 & !is.na(data$Cumulative_deaths),]$Cumulative_deaths))

```

```{r}
matrice_data <- matrix(data$Cumulative_deaths, byrow=TRUE, ncol=1238)

liste_dedoublee <- unlist(replicate(1238,liste_pays, simplify = FALSE), recursive = FALSE)
liste_dedoublee <- matrix(liste_dedoublee, byrow=FALSE, ncol=30)
dataset <- data.frame(time = rep(1:1238,30),
                     country = as.vector(t(matrix(liste_dedoublee, byrow=FALSE, ncol=30))),
                     cumulative_count = as.vector(t(matrice_data)))
```

```{r}
max <- max(data$Cumulative_deaths)
min <- min(data$Cumulative_deaths)
plot(1:1238, colMeans(matrice_data),type="l", ylim=c(min, max), col="red")
for (country in 1:30){
  points(1:1238, matrice_data[country,],type="l")
}
points(1:1238, colMeans(matrice_data),type="l", ylim=c(min, max), col="red")
```



```{r}
perform_pca <- function(X){
  mu <- colMeans(X)
  X <- sweep(X,2,mu)
  
  SVD <- svd(X)
  Scores <- SVD$u %*% diag(SVD$d)
  Loadings <- SVD$v
  FVE <- SVD$d^2/sum(SVD$d^2)
  
  lam <- sqrt(length(xeval)) # measure change
  op <- par(mfrow=c(3,2),mar=rep(2,4))
  plot(xeval, X[1,]+mu,type="l", ylim=range(X+mu), main="Data and the mean")
  for(n in 1:dim(X)[1]) points(xeval, X[n,]+mu,type="l")
  points(xeval,mu,col=2,lwd=2,type="l")
  plot(Scores[1,]*sign(sum(Loadings[,1])), Scores[2,]*sign(sum(Loadings[,2])), main="1st vs 2nd PC scores")
  
  plot(xeval,Loadings[,1]*sign(sum(Loadings[,1])),type="l", main=paste0("1st PC (",round(100*FVE[1])," % of var)"))
  plot(xeval,Loadings[,2]*sign(sum(Loadings[,2])),type="l", main=paste0("2nd PC (",round(100*FVE[2])," % of var)"))
  plot(xeval,Loadings[,3]*sign(sum(Loadings[,3])),type="l", main=paste0("3rd PC (",round(100*FVE[3])," % of var)"))
  plot(xeval,Loadings[,4]*sign(sum(Loadings[,4])),type="l", main=paste0("4th PC (",round(100*FVE[4])," % of var)"))
}

xeval <- seq(1, 1238 ,by=1)
perform_pca(matrice_data)
``` 



# Cut and smoothing 

Before applying our PCA, we decide to apply a smoothing method to each of the curves corresponding to a country. This reduces the variability of the curves and makes them smoother. In fact, we have discrete data (each day) whereas in reality we're studying a continuous phenomenon. Noisy data can indeed be a problem for PCA, so these methods enable better structure identification. 

[Utilisation de la méthode du kernel]

```{r}
matrice_cut <- matrix(0, nrow=30, ncol=100)
for (i in 1:30){
  indice <- which.min(abs(matrice_data[i,] +1 ))
  matrice_cut[i,] <- matrice_data[i,indice:(indice+99)]
}
```

```{r}
matrice_cut_smooth <- matrice_cut
for (i in 1:30){
  matrice_cut_smooth[i,] <- ksmooth(1:121, matrice_cut_smooth[i,], "normal", bandwidth = 10)$y
}
```


```{r}
max <- max(matrice_cut_smooth)
min <- min(matrice_cut_smooth)
plot(1:100, colMeans(matrice_cut_smooth),type="l", ylim=c(min, max), col="red")
for (country in 1:30){
  points(1:100, matrice_cut[country,],type="l")
}
```

```{r}
xeval <- seq(1, 100 ,by=1)
perform_pca(matrice_cut)
```

